### 2018-01-31 includes/abstract.tex
Increasing consistency. Renaming problems.tex to suites.tex. Using the word 'instance' instead of the word 'problem'.

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem
    transformer and generator for string solvers. We supply a
    repository of problems generated by \fuzzer{} in \smtfull{}
    format. We systematically compare \theSolvers{} on groups of such
    problems, and identify problems that are particularly challenging for some
    solvers. We briefly explain our observations and show how
    \fuzzer{} helped identify causes of performance degradations in
    \us{}.
\end{abstract}

>>>>

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem instance
    transformer and generator for string solvers. We supply a
    repository of instances generated by \fuzzer{} in \smtfull{}
    format. We systematically compare \theSolvers{} on groups of such
    instances, and identify instances that are particularly challenging for
    some solvers. We briefly explain our observations and show how
    \fuzzer{} helped identify causes of performance degradations in
    \us{}.
\end{abstract}

### 2018-01-31 includes/experiment.tex
Increasing consistency. Renaming problems.tex to suites.tex. Using the word 'instance' instead of the word 'problem'.

We generated several problem suites with \fuzzer{} that made one
solver perform poorly, but not others. These suites are
\theSuites{}. Figure~\ref{fig:cvc-hard} shows the suites that were
uniquely difficult for \cvc{}. Figure~\ref{fig:z3str3-hard} shows the
suites that were uniquely difficult for \us{}. All experiments were
run in series, with a timeout of 15 seconds, on the same computer running
Ubuntu Linux 16.04. The computer had 32GB of RAM and an
Intel\textregistered{} Core\texttrademark{} i7-6700 CPU with clock speed
of 3.40GHz.

\begin{figure}[h]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-extracts-small.png}
        \caption{Performance on concats-extracts-small}
        \label{fig:concats-extracts-small}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/different-prefix.png}
        \caption{Performance on different-prefix}
        \label{fig:different-prefix}
    \end{subfigure}
    \caption{Problems hard for \cvc{}}
    \label{fig:cvc-hard}

    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-balanced.png}
        \label{fig:concats-balanced}
        \caption{Performance on concats-balanced}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-small.png}
        \label{fig:concats-small}
        \caption{Performance on concats-small}
    \end{subfigure}
    \caption{Problems hard for \us{}}
    \label{fig:z3str3-hard}
\end{figure}

@@

We found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \textit{concats-big} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As an immediate consequence of
testing \us{} on \textit{concats-big} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly.

>>>>

We generated several problem instance suites with \fuzzer{} that made one
solver perform poorly, but not others. These suites are
\theSuites{}. Figure~\ref{fig:cvc-hard} shows the suites that were
uniquely difficult for \cvc{}. Figure~\ref{fig:z3str3-hard} shows the
suites that were uniquely difficult for \us{}. All experiments were
run in series, with a timeout of 15 seconds, on the same computer running
Ubuntu Linux 16.04. The computer had 32GB of RAM and an
Intel\textregistered{} Core\texttrademark{} i7-6700 CPU with clock speed
of 3.40GHz.

\begin{figure}[h]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-extracts-small.png}
        \caption{Performance on concats-extracts-small}
        \label{fig:concats-extracts-small}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/different-prefix.png}
        \caption{Performance on different-prefix}
        \label{fig:different-prefix}
    \end{subfigure}
    \caption{Instances hard for \cvc{}}
    \label{fig:cvc-hard}

    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-balanced.png}
        \label{fig:concats-balanced}
        \caption{Performance on concats-balanced}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-small.png}
        \label{fig:concats-small}
        \caption{Performance on concats-small}
    \end{subfigure}
    \caption{Instances hard for \us{}}
    \label{fig:z3str3-hard}
\end{figure}

@@

\fuzzer{}'s ability to produce scaling instances (e.g., long string constants,
concats of increasing depth, etc.) helped uncover several
implementation issues and performance limitations in \us{}. Scaling inputs
can reveal issues that would be out of scope for unit tests or
industrial benchmarks because they elicit asymptotic behaviour. Three different
performance and implementation bugs were identified and fixed in \us{}
as a result of testing with the \fuzzer{} suites
\textit{Lengths} and \textit{Concats}.

We also found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \textit{Concats} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As an immediate consequence of
testing \us{} on \textit{Concats} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly.

### 2018-01-31 includes/introduction.tex
Increasing consistency. Renaming problems.tex to suites.tex. Using the word 'instance' instead of the word 'problem'.

\begin{enumerate}
    \item \textbf{The \fuzzer{} tool}:
        In section~\ref{sec:fuzzer}, we describe a modular fuzzer that can
        transform and randomly generate \smtfull{} string and regular
        expression (regex) problems.\footnote{We assume basic
        familiarity with string solvers and their input
        language.} SScaling inputs (e.g., long string constants,
        deep concatenations) are particularly useful in identifying asymptotic
        behaviors in solvers, and \fuzzer{} has many options to generate such
        scaling inputs. We briefly document \fuzzer{}'s
        components and modular architecture. We provide example use cases to
        demonstrate its utility as an exploratory tool for validating
        solvers.

    \item \textbf{A repository of \smtfull{} problems}:
        We present a repository of \smtfull{} string and regex problem suites
        we generated using \fuzzer{} in section~\ref{sec:problems}. The
        repository consists of two categories: \texttt{generated}, with new
        problems generated by \fuzzer{}, and \texttt{transformed}, with
        transformed problems generated from a small suite of industrial
        benchmarks.

>>>>

\begin{enumerate}
    \item \textbf{The \fuzzer{} tool}:
        In section~\ref{sec:fuzzer}, we describe a modular fuzzer that can
        transform and randomly generate \smtfull{} string and regular
        expression (regex) instances.\footnote{We assume basic
        familiarity with string solvers and their input
        language.} Scaling inputs (e.g., long string constants,
        deep concatenations) are particularly useful in identifying asymptotic
        behaviors in solvers, and \fuzzer{} has many options to generate such
        scaling inputs. We briefly document \fuzzer{}'s
        components and modular architecture. We provide example use cases to
        demonstrate its utility as an exploratory tool for validating
        solvers.

    \item \textbf{A repository of \smtfull{} instances}:
        We present a repository of \smtfull{} string and regex problem suites
        we generated using \fuzzer{} in section~\ref{sec:suites}. The
        repository consists of two categories: \texttt{generated}, with new
        instances generated by \fuzzer{}, and \texttt{transformed}, with
        transformed instances generated from a small suite of industrial
        benchmarks.

### 2018-01-31 includes/prior.tex
cleaned up the paper a lot
First pass with Federico.
Increasing consistency. Renaming problems.tex to suites.tex. Using the word 'instance' instead of the word 'problem'.

    Naturally, many solver developers author their own test suites to validate their solvers~\cite{cvc4-tests,z3str3-tests,z3str2-tests}. In addition, several popular problem suites are publicly available for solver validation, such as the Kaluza~\cite{kaluza} and Kausler~\cite{kausler} suites.

    There are likewise several fuzzers and problem generators currently available, but none of them can generate or transform string and regex problems. For example, the FuzzSMT~\cite{fuzzsmt} tool generates \smt{} problems with bit-vectors and arrays, but does not support strings or regexes. The SMTpp~\cite{smtpp} tool pre-processes and simplifies problems, but does not generate new ones or fuzz existing ones.

% FuzzSMT: http://fmv.jku.at/papers/BrummayerBiere-SMT09.pdf
% SMTpp: http://www.verit-solver.org/papers/smt2015.pdf

>>>>

Many solver developers author their own test suites to
validate their solvers~\cite{cvc4-tests,z3str3-tests,z3str2-tests}. In
addition, several popular instance suites are publicly available for
solver testing and benchmarking, such as the Kaluza~\cite{kaluza} and
Kausler~\cite{kausler} suites. There are likewise several fuzzers and
instance generators currently available, but none of them can generate
or transform string and regex instances. For example, the
FuzzSMT~\cite{fuzzsmt} tool generates \smt{} instances with bit-vectors
and arrays, but does not support strings or regexes. The
SMTpp~\cite{smtpp} tool pre-processes and simplifies instances, but
does not generate new ones or fuzz existing ones.

### 2018-01-31 includes/stringfuzz.tex
Increasing consistency. Renaming problems.tex to suites.tex. Using the word 'instance' instead of the word 'problem'.

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
problems. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{translate} and \textit{reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[t]
    \centering
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{concats}     & Long concats and optional random extracts. \\ \hline
        \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{regex}       & Regexes of varying complexity. \\ \hline
        \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{nop}         & Does nothing. \\ \hline
        \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

\fuzzer{} can also transform the instances it generates. All of the
transformers in \transformer{} can operate on regex terms. They can
multiply and permute regex literals (e.g. \textit{reverse},
\textit{multiply}, \textit{translate}), randomly replace regex
operators (e.g. \textit{fuzz}), and manipulate the syntax tree
(e.g. \textit{graft}, \textit{rotate}).

>>>>

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
problems. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{Translate} and \textit{Reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[t]
    \centering
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{Concats}     & Long concats and optional random extracts. \\ \hline
        \textit{Lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{Overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{Equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{Regex}       & Regexes of varying complexity. \\ \hline
        \textit{Random-Text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{Random-AST}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{Fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{Graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{Multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{Nop}         & Doesn't transform (but can translate \smtfull{}). \\ \hline
        \textit{Reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{Rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{Translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{Unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

\fuzzer{} can also transform the instances it generates. All of the
transformers in \transformer{} can operate on regex terms. They can
multiply and permute regex literals (e.g. \textit{Reverse},
\textit{Multiply}, \textit{Translate}), randomly replace regex
operators (e.g. \textit{Fuzz}), and manipulate the syntax tree
(e.g. \textit{Graft}, \textit{Rotate}).
